---
title: "January 27"
output: 
  bookdown::pdf_document2:
    number_sections: false
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE, dev = "ragg_png", dpi=200)
library(tidyverse)
library(patchwork)

source(here::here("R", "updates.R"))
# Sys.setenv(TAR_PROJECT = "ale")
# tar_config_set(store = here::here("_ale"))

point_size <- 1

plot_dist <- function(outqs, min_cluster=10, ps=point_size){
  qs::qread(outqs) |>
    filter(!str_detect(label, "Brain-Stem")) |>
    filter(cluster_size > min_cluster) |>
    # mutate(label = str_c(label, cluster_size, sep = " ")) |>
    group_by(label, n_sub, iter, sign) |>
    summarise(d = min(d), .groups = "drop") |>
    mutate(n_sub = factor(n_sub), sign=factor(sign)) |>
    # WISEsummary(
    #   dependentvars = "d",
    #   betweenvars = c("n_sub", "label", "cluster_size","sign")) |>
    ggplot(aes(y=n_sub, x=d)) + 
    facet_wrap(
      ~label, 
      # scales = "free_y",
      labeller = ggplot2::label_wrap_gen(width = 20)) +
    ggridges::geom_density_ridges(aes(colour=sign),
                                  jittered_points = TRUE,
                                  position = ggridges::position_points_jitter(width = 0.05, height = 0),
                                  point_shape = '|', 
                                  point_size = ps, 
                                  point_alpha = 1, 
                                  alpha = 0.7,
                                  rel_min_height = 0.01) +
    # geom_line() +
    # scale_x_continuous(trans = "log10") +
    # geom_errorbar(aes(ymin=d_CI_lower, ymax=d_CI_upper)) +
    xlab("Avg. Distance From Reference (mm)") +
    ylab("N Participants") +
    theme_gray(base_size = 8) 
}

plot_tmp <- function(outqs, min_cluster=10){
  qs::qread(outqs) |>
    filter(!str_detect(label, "Brain-Stem")) |>
    filter(cluster_size > min_cluster) |>
    mutate(label = str_c(label, sep = " ")) |>
    group_by(label, n_sub, iter, sign) |> 
    filter(d == min(d)) |> 
    ungroup() |> 
    group_nest(n_sub, label, sign) |> 
    mutate(
      avg = map(
        data, 
        ~tibble(d=as.vector(dist(cbind(.x$x.study,.x$y.study,.x$z.study)))) |> 
          WISEsummary(dependentvars = "d") )) |> 
    select(-data) |> 
    unnest(avg) |> 
    ggplot(aes(x=n_sub, y=d_mean, color=sign)) + 
    facet_wrap(
      ~label, 
      labeller = ggplot2::label_wrap_gen(width = 20)) +
    geom_line() +
    scale_x_continuous(trans = "log10") +
    geom_errorbar(aes(ymin=d_CI_lower, ymax=d_CI_upper)) +
    ylab("Avg. Distance Between Studies (mm)") +
    xlab("N Participants") +
    theme_gray(base_size = 8)
}

```


## Biobank

I encountered a few more participants that appear to have been downloaded incompletely (3694178, 3696599, and 3702736). Of course, there are still plenty to process. But is there a good place to record these?

## spacetop

As you've said, a couple of terabytes have opened up on `/dcl01/smart/data`. That space is quickly being eaten up by processing these data. Much of that is from the intermediate files produced by MRIQC and fMRIPrep (e.g., the raw data is a bit over 1T, and these pipelines make at a couple of copies of the raw data files each). I'll see this afternoon whether they want to store the intermediate files. So, this is partly to explain why space might be limited again, if anyone asks about that in the near-term.

## Population Effect Sice / Precision

The story about worsening localization of local peaks ended up being simple. There were two issues. The first is largely one of me not understanding the default settings for FSL's `cluster`. It calculates local maxima/minima through a straightforward strategy: a voxel is a local peak if it is inside a cluster and it has a higher value than all of its neighbors (it allows for either 6 or 26 neighbors). Given that, it was very strange that the large clusters visible in studies of 100 + participants were not producing local peaks that matched the local peaks in the reference study. The issue is that, by default, `cluster` limits the number of local peaks within each cluster to six, the six with the largest value. That limit may not be reached when clusters are isolated, but at larger sample sizes the clusters are very large and have many than six local peaks.

Figures 1 and 2 better portray how increasing sample sizes affects peak localization. 

```{r, space, fig.cap="Average Distance of Study Peaks From Reference. Panel labels correspond to the location of the local peak in the reference study, which could be a maximum or minimum. Cluster forming threshold was p < 0.001.", fig.width=9, fig.height=9}

plot_dist(here::here("data-raw/out3.1.qs")) +
  xlim(0,40)

```



```{r, precision, fig.cap="Average Distance of Study Peaks From Each Other. Panel labels correspond to the location of the local peak in the reference study. Cluster forming threshold was p < 0.001. Error bars span 95% confidence intervals.", fig.width=9, fig.height=9}

plot_tmp(here::here("data-raw/out3.1.qs"))

```

There was a second issue, indicated by how localization of the Hippocampi still appears to worsen as sample size increases. This is caused by overflow in the z-maps, such that the voxels with the highest activity all have values of `inf`. This started with the more recent increases in the size of the reference study. The local peaks are defined with the rule that `inf == inf`, so the "local peaks" end up on the edges of the overflow region. At the current sample size, overflow is pronounced in the Hippocampi. Whether individual studies have peaks that are close to the reference study depends on whether the individual study also has overflow, and whether the overflowing voxels match those in the reference study.

```{r, spaceC, fig.cap="Peak localization impeded by numerical overflow. A) Localization of right hippocampus appears to worsen with larger studies. B) Thresholded z-map from reference study. Voxels with overflowing values are colored brown. Cross-hairs placed on 'local peak', spuriously chosen.", fig.width=9, fig.height=9}

out <- qs::qread(here::here("data-raw/out3.1.qs")) |>
  filter(!str_detect(label, "Brain-Stem")) |>
  filter(cluster_size > 10) |>
  mutate(label = str_c(label, sep = " ")) |>
  group_by(label, n_sub, iter) |>
  slice_min(d, n=1) |>
  ungroup() |>
  mutate(n_sub = factor(n_sub)) 

label.study <- out |> 
  distinct(label, label.study) |>
  group_nest(label) |>
  mutate(
    data = map(
      data, 
      ~distinct(.x,label.study) |>
        mutate(ls = factor(label.study) |> as.numeric()))) |>
  unnest(data) |>
  mutate(ls = factor(ls))

a <- out |>
  filter(str_detect(label, "Hipp")) |>
  left_join(label.study, by = c("label","label.study")) |>
  mutate(
    label.hemi = if_else(x.study > 45, "Left ", "Right "),
    label.study = if_else(str_detect(label.study, "Left|Right"), label.study, str_c(label.hemi, label.study))) |>
  ggplot(aes(y=n_sub, x=d)) + 
  facet_wrap(
    ~label, 
    # scales = "free_y",
    labeller = ggplot2::label_wrap_gen(width = 20)) +
  ggridges::geom_density_ridges(
    alpha = 0.7,
    rel_min_height = 0.01) +
  geom_point(aes(color=label.study), shape = 3, stroke=3, alpha =0.3) + 
  xlab("Avg. Distance From Reference (mm)") +
  ylab("N Participants") +
  theme_gray(base_size = 8) 

b <- patchwork::wrap_elements(
  ~imager::load.image(here::here("analysis","updates","img", "overflow.png")) |>
  imager:::plot.cimg(axes = FALSE))

a / b + plot_annotation(tag_levels = "A", tag_suffix = ")")

```



